### 2015-02-18 fundamentals/conclusion.tex
tiny changes
spell check and trailing whitespaces
the word code is not appropriate when we mean software or simulator. that is, it doesn't make the best noun.

The \Cyclus nuclear fuel cycle framework presents a more generic and flexible 
alternative to existing fuel cycle simulators.

Where previous nuclear fuel cycle simulators have had limited distribution, 
uncertain provenance, and restricted customizability, \Cyclus emphasizes an 
open access and development strategy.  Open development, with transparent 
publication of the source code, not only provides much-needed accessibility, 
but also enables transparency and community oversight.

The object-oriented \acrlong{ABM} simulation paradigm ensures more generic 
simulation capability. It allows \Cyclus to address common analyses in a more 
flexible fashion and enables analyses that are impossible with system dynamics 
codes.

>>>>

The \Cyclus nuclear fuel cycle framework presents a more generic and flexible
alternative to existing fuel cycle simulators. Where previous nuclear fuel cycle simulators have had limited distribution,
uncertain provenance, and restricted customizability, \Cyclus emphasizes an
open access and development strategy.  Open development, with transparent
publication of the source code, not only provides much-needed accessibility,
but also enables transparency and community oversight.

The object-oriented \gls{ABM} simulation paradigm ensures more generic
simulation capability. It allows \Cyclus to address common analyses in a more
flexible fashion and enables analyses that are impossible with system dynamics
simulators.

### 2015-02-18 fundamentals/fcsreqs.tex
the word code is not appropriate when we mean software or simulator. that is, it doesn't make the best noun.

The proprietary concerns of research institutions and security constraints of
data within fuel cycle codes often restrict code access. A code's use is
therefore often limited to its institution of origin, necessitating effort
duplication at other institutions and thereby squandering broader human
resources. License agreements and institutional approval are required for most
current codes (i.e. \gls{COSI}6, \gls{DANESS}, \gls{DESAE}, EVOLCODE,
FAMILY21, \gls{NFCSim})\cite{juchau_modeling_2010}, ORION, and VISION.  Even when, as in
the case of the MIT \gls{CAFCA} software, the source code is unrestricted, the
platform on which it relies is often restricted or commercial.  However, \Cyclus
provide fully free and open access to all users
and developers, foreign and domestic.

>>>>

The proprietary concerns of research institutions and security constraints of
data within fuel cycle simulators often restrict access. Use of a simulator is
therefore often limited to its institution of origin, necessitating effort
duplication at other institutions and thereby squandering broader human
resources. License agreements and institutional approval are required for most
current simulators (i.e. \gls{COSI}6, \gls{DANESS}, \gls{DESAE}, EVOLCODE,
FAMILY21, \gls{NFCSim})\cite{juchau_modeling_2010}, ORION, and VISION.  Even when, as in
the case of the MIT \gls{CAFCA} software, the source code is unrestricted, the
platform on which it relies is often restricted or commercial.  However, \Cyclus
provides fully free and open access to all users
and developers, foreign and domestic.

### 2015-02-18 fundamentals/impl.tex
the word code is not appropriate when we mean software or simulator. that is, it doesn't make the best noun.

The architecture of \Cyclus allows developers
to define nuclear fuel cycle agents and processes independent of the simulation
logic. To achieve this, agents are developed using the \Cyclus framework \gls{API}, a set of functions and protocols
which
assist in agent development and specify how agents should be defined.
This encapsulated `plug-in'
design choice provides two major benefits. First, analysts across research institutions can take advantage of the
simulation logic \gls{API} and archetype ecosystem when they apply \Cyclus to
their specific problem. This
mend-and-extend paradigm, common in modern software development, reduces
maintenance overhead and improves code robustness by protecting the core logic
from changes in the extensions.
The researcher can focus on creating or customizing nuclear
facility, institution, resource, and toolkit models within their specific area
of technical expertise. This extensibility, magnified by an open source
development process, frees the community of users to contribute desired
features, independent from the team of \Cyclus core developers.
Second, because \Cyclus uses a modular archetype approach, comparing two
archetypes is straightforward. For example, if an analyst would like
to compare the effect of using different models to determine the input fuel
composition for fast reactors, a variety of fuel fabrication archetypes can be
developed and interchanged while keeping the rest of the fuel cycle fixed
 to compare fuel fabrication designs.

@@

\Cyclus acheives this behavior by implementing its generic \gls{API} and
modular architecture via a suite of dynamically loadable plug-in libraries.
Though common in modern software architecture, such a plug-in pattern has not
previously been implemented in nuclear fuel cycle simulators.  Dynamically
loadable libraries are the primary mechanism for extending \Cyclus' capability.
This approach provides encapsulation: the core of the code operates
independently from the individual plug-in libraries. Thus, any customization or
extension is implemented only in the loadable library.

@@

Nuclear engineering code quality is often governed by \gls{NQA1}, an
\gls{ASME} specification
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}.
Although \gls{NQA1} is primarily
used for designing reactors, it is general enough to apply to
\Cyclus. \Cyclus has adopted an \emph{agile} development process
\cite{larman_agile_2004},
interpreting \gls{NQA1} in a manner similar to the process adopted by the
\gls{DOE} within \gls{NEAMS} \cite{neams_nuclear_2013} or the
PyNE toolkit \cite{biondo_quality_2014}.

Quality assurance is an on-going process throughout the
entire life of the code. As a simulator where most modeling decisions are made
by third party archetype developers or users, the most important feature of
\gls{QA} is verification. Validation may be impossible and \gls{UQ} is beyond
the scope of the kernel.

@@

All of the public interface (the \gls{API}) must be documented as required by the \Cyclus
\gls{QA} policy. This certifies that the intention for a code unit is communicated
in way that is separate from its implementation. This policy is important
because implementations are often wrong and do not reliably communicate their intent. Truly trustworthy communication about what
software should be doing occurs when the author explicitly states the
goals of an \gls{API} in prose form. Future development of the code is then able to
compare the implementation versus the documentation to discern whether or not they
are in agreement. Documentation is thus a secondary higher-level information
stream.  In \Cyclus, this information is aggregated together into static
websites with the Doxygen \cite{van_heesch_doxygen:_2008} and Sphinx
\cite{brandl_sphinx_2014}
tools, and can be accessed at \url{http://fuelcycle.org}.

@@

Version control is the history-preserving mechanism by which the provenance of
a code is recorded. \Cyclus uses a well-established tool called git
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utility in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
Though there are many praises and criticisms of git, one
feature that the tool performs exceptionally well is the notion of code branches.
These are distinct pathways in the history that branch out of a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simultaneous
branches may exist at all points in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an
accompanying message. Thus
it is possible for \Cyclus to accurately replay the entirety of who did what to the
code when.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[height=0.6\textheight]{./images/gitprocess}
\end{center}
\caption{A version-control enabled process for finding, reporting, and fixing a bug
can take place within an issue tracker alongside multiple stages of code
testing and review. In the \Cyclus case, the final review stage includes unit,
regression, and integration testing as well as manual code inspection.}
\label{fig:gitprocess}
\end{figure}

>>>>

The architecture of \Cyclus allows developers
to define nuclear fuel cycle agents and processes independent of the simulation
logic. To achieve this, agents are developed using the \Cyclus framework \gls{API}, a set of functions and protocols
which
assist in agent development and specify how agents should be defined.
This encapsulated `plug-in'
design choice provides two major benefits. First, analysts across research institutions can take advantage of the
simulation logic \gls{API} and archetype ecosystem when they apply \Cyclus to
their specific problem. This
mend-and-extend paradigm, common in modern software development, reduces
maintenance overhead and improves software robustness by protecting the core logic
from changes in the extensions.
The researcher can focus on creating or customizing nuclear
facility, institution, resource, and toolkit models within their specific area
of technical expertise. This extensibility, magnified by an open source
development process, frees the community of users to contribute desired
features, independent from the team of \Cyclus core developers.
Second, because \Cyclus uses a modular archetype approach, comparing two
archetypes is straightforward. For example, if an analyst would like
to compare the effect of using different models to determine the input fuel
composition for fast reactors, a variety of fuel fabrication archetypes can be
developed and interchanged while keeping the rest of the fuel cycle fixed
 to compare fuel fabrication designs.

@@

\Cyclus acheives this behavior by implementing its generic \gls{API} and
modular architecture via a suite of dynamically loadable plug-in libraries.
Though common in modern software architecture, such a plug-in pattern has not
previously been implemented in nuclear fuel cycle simulators.  Dynamically
loadable libraries are the primary mechanism for extending \Cyclus' capability.
This approach provides encapsulation: the core framework operates
independently from the individual plug-in libraries. Thus, any customization or
extension is implemented only in the loadable library.

@@

Nuclear engineering software quality is often governed by \gls{NQA1}, an
\gls{ASME} specification
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}.
Although \gls{NQA1} is primarily
used for designing reactors, it is general enough to apply to
\Cyclus. \Cyclus has adopted an \emph{agile} development process
\cite{larman_agile_2004},
interpreting \gls{NQA1} in a manner similar to the process adopted by the
\gls{DOE} within \gls{NEAMS} \cite{neams_nuclear_2013} or the
PyNE toolkit \cite{biondo_quality_2014}.

Quality assurance is an on-going process throughout the
entire life of the software. As a simulator where most modeling decisions are made
by third party archetype developers or users, the most important feature of
\gls{QA} is verification. Validation may be impossible and \gls{UQ} is beyond
the scope of the kernel.

@@

All of the public interface (the \gls{API}) must be documented as required by the \Cyclus
\gls{QA} policy. This certifies that the intention for a code unit is communicated
in way that is separate from its implementation. This policy is important
because implementations are often wrong and do not reliably communicate their intent. Truly trustworthy communication about what
software should be doing occurs when the author explicitly states the
goals of an \gls{API} in prose form. Future developers are then able to
compare the implementation versus the documentation to discern whether or not
these are in agreement. Documentation is thus a secondary higher-level information
stream.  In \Cyclus, this information is aggregated together into static
websites with the Doxygen \cite{van_heesch_doxygen:_2008} and Sphinx
\cite{brandl_sphinx_2014}
tools, and can be accessed at \url{http://fuelcycle.org}.

@@

Version control preserves the development history, or \emph{provenance}, of
source code files. \Cyclus uses a well-established tool called git
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utility in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
One feature that git performs exceptionally well is \emph{branching}.
Branches are distinct pathways in the history that diverge from a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simultaneous
branches may exist at all points in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an
accompanying message. Thus
it is possible for \Cyclus to accurately replay the entirety of who did what to the
code when.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[height=0.6\textheight]{./images/gitprocess}
\end{center}
\caption{A version-control enabled process for finding, reporting, and fixing a bug
can take place within an issue tracker alongside multiple stages of software
testing and review. In the \Cyclus case, the final review stage includes unit,
regression, and integration testing as well as manual code inspection.}
\label{fig:gitprocess}
\end{figure}

### 2015-02-18 fundamentals/intro.tex
the word code is not appropriate when we mean software or simulator. that is, it doesn't make the best noun.

\Cyclus, therefore, is a dynamic, agent-based model, which employs a modular
architecture, an open development process, discrete agents, discrete time, and
arbitrarily detailed isotopic resolution of materials. Dynamic nuclear fuel
cycle analysis more realistically supports a range of simulation goals than
static analysis \cite{piet_dynamic_2011}. Similarly, experience in the broader
field of systems analysis indicates that agent-based modeling enables more
flexible simulation control, without loss of generality
\cite{macal_agent-based_2010}. Finally, openness allows cross-institutional
collaboration, increases code robustness \cite{cohen_modern_2010}, and
cultivates an ecosystem of modeling and calculation options.  Comprising
flexible, interchangeable models of fuel cycle components with varying scope,
depth and fidelity, this ecosystem allows users and developers to customise
\Cyclus to analyze the cases that are of interest to them, rather than those
the simulator was originally developed to address.  The fundamental concepts of
the \Cyclus nuclear fuel cycle simulator capture these modern insights so that
challenges in nuclear fuel cycle analysis can be better addressed.

@@

However, methods for calculating those metrics vary among simulators. Some
model the system of facilities, economics, and materials in static equilibrium,
while other simulators capture the dynamics of the system.  Similarly, while
some simulators discretely model batches of material and individual facilities,
others aggregate facilities into fleets and materials into streams. Simulators
can model a single aspect of the fuel cycle in great detail while neglecting
others. For example, a simulator created for policy modeling might have
excellent capability in economics while models affecting transformations in
material isotopics or the effects of isotopics on technology performance are
neglected.  The \gls{CAFCA}\cite{guerin_impact_2009} code is problem-oriented
in this way.

@@

\Cyclus emerged from a line of tools seeking to break this practice.  Its
precursor,
\gls{GENIUSv1}\cite{dunzik-gougar_global_2007,jain_transitioning_2006},
originated within \gls{INL} and sought to provide generic regional capability.
Based on lessons learned from Version 1, the \acrshort{GENIUSv2} Version
2\cite{oliver_studying_2009,huff_geniusv2_2009} code sought to provide more
generality and an extensible interface to facilitate collaboration.  The
\Cyclus project then improved upon the \acrshort{GENIUSv2} effort by
implementing increased modularity and encapsulation.  The result is a dynamic
simulator that treats both materials and facilities discretely, with an
architecture that permits multiple and variable levels of fidelity. The
simulator's agent-based framework for tracking the transformation and trade of
resources between autonomous regional and institutional entities with
customizable behavior and objectives is an innovation not pursued by any
existing fuel cycle simulator.

>>>>

\Cyclus, therefore, is a dynamic, agent-based model, which employs a modular
architecture, an open development process, discrete agents, discrete time, and
arbitrarily detailed isotopic resolution of materials. Dynamic nuclear fuel
cycle analysis more realistically supports a range of simulation goals than
static analysis \cite{piet_dynamic_2011}. Similarly, experience in the broader
field of systems analysis indicates that agent-based modeling enables more
flexible simulation control, without loss of generality
\cite{macal_agent-based_2010}. Finally, openness allows cross-institutional
collaboration, increases software robustness \cite{cohen_modern_2010}, and
cultivates an ecosystem of modeling and calculation options.  Comprising
flexible, interchangeable models of fuel cycle components with varying scope,
depth and fidelity, this ecosystem allows users and developers to customise
\Cyclus to analyze the cases that are of interest to them, rather than those
the simulator was originally developed to address.  The fundamental concepts of
the \Cyclus nuclear fuel cycle simulator capture these modern insights so that
challenges in nuclear fuel cycle analysis can be better addressed.

@@

However, methods for calculating those metrics vary among simulators. Some
model the system of facilities, economics, and materials in static equilibrium,
while other simulators capture the dynamics of the system.  Similarly, while
some simulators discretely model batches of material and individual facilities,
others aggregate facilities into fleets and materials into streams. Simulators
can model a single aspect of the fuel cycle in great detail while neglecting
others. For example, a simulator created for policy modeling might have
excellent capability in economics while models affecting transformations in
material isotopics or the effects of isotopics on technology performance are
neglected.  The \gls{CAFCA}\cite{guerin_impact_2009} simulator is problem-oriented
in this way.

@@

\Cyclus emerged from a line of tools seeking to break this practice.  Its
precursor,
\gls{GENIUSv1}\cite{dunzik-gougar_global_2007,jain_transitioning_2006},
originated within \gls{INL} and sought to provide generic regional capability.
Based on lessons learned from Version 1, the \acrshort{GENIUSv2} Version
2\cite{oliver_studying_2009,huff_geniusv2_2009} simulator sought to provide more
generality and an extensible interface to facilitate collaboration.  The
\Cyclus project then improved upon the \acrshort{GENIUSv2} effort by
implementing increased modularity and encapsulation.  The result is a dynamic
simulator that treats both materials and facilities discretely, with an
architecture that permits multiple and variable levels of fidelity. The
simulator's agent-based framework for tracking the transformation and trade of
resources between autonomous regional and institutional entities with
customizable behavior and objectives is an innovation not pursued by any
existing fuel cycle simulator.
